deckard: A Declarative Tool for Machine Learning Robustness Evaluations

### 2025-05-09 paper.tex
Commit from GitHub Actions (Render R Markdown)
typos

Various versions of this software have been used in several recently
published and not-yet-published works by the author of this paper, all
of which are available in the \texttt{examples} folder in the source
code repository \url{https://github.com/simplymathematics/deckard}. One
published work, now reproducible via the
\texttt{examples/attack\_defence\_survey} folder, includes a large
survey of attacks and defences against canonical datasets and models (C.
Meyers, Löfstedt, and Elmroth 2023). Another work analysed the run-time
requirements of attacks against a particular model before and after
retraining against those attacks (Charles Meyers and Elmroth 2024)
(reproducible via \texttt{examples/retraining}). The next paper
formalised a method for estimating the time-to-failure of a given model
against a suite of attacks and introduce a metric that quantifies the
ratio of attack and training cost (Meyers et al. 2023) (reproducible via
\texttt{examples/survival\_heuristic}). Furthermore, a not yet published
work uses this time-to-failure model as a mechanism for analysing the
cost efficacy of various hardware choices in the context of adversarial
attacks (reproducible via \texttt{examples/power}) (C. Meyers et al.
2024). Another work exploits the tooling to train a custom model that is
designed to run client-side by using compression algorithms to measure
the distance between text (reproducible via
\texttt{examples/compression}).

@@

Since ML projects can exploit specialized hardware such as multi-core
processors or GPUs, and often rely on clusters of machines for
large-scale data processing, it was necessary to enable parallel and
distributed experiment execution and model optimization. By leveraging
the \texttt{hydra} configuration framework, \texttt{deckard}
automatically supports optimization libraries like \texttt{nevergrad}
(Bennet et al. 2021), \texttt{Adaptive\ Experimentation} (A. Developers
2025), and \texttt{optuna} (Akiba et al. 2019), making the software
modular and extensible. Additionally, experiments can be managed using a
variety of popular job schedulers, including \texttt{Ray} (Moritz et al.
2018), \texttt{Redis\ Queue} (Stamps 2025), and \texttt{slurm} (Yoo,
Jette, and Grondona 2003) for distributd jobs or \texttt{joblib} (Joblib
Developers 2025) for jobs on a single machine.

By using a declarative design, a given set of experiments can be
specified once and executed seamlessly across different backends without
modification to the underlying codebase. This makes \texttt{deckard}
both adaptable and scalable, suitable for use on personal laptops,
multi-node servers, or large-scale, high-performance clusters. When
configured appropriately, experiment batches can be parallelized,
enabling massive parameter sweeps, ensemble evaluations, or adversarial
robustness tests to be executed in parallel---reducing turnaround time
while maintaining strong guarantees on reproducibility and auditability.
The design of the presented software prioritizes clarity and
maintainability by capturing each experimental configuration as a YAML
artifact, making both successful and failed runs equally traceable and
shareable. This approach transforms experiment tracking from an
afterthought into a first-class component of the trustworthy ML
workflow.

>>>>

Various versions of this software have been used in several recently
published and not-yet-published works by the author of this paper, all
of which are available in the \texttt{examples} folder in the source
code repository \url{https://github.com/simplymathematics/deckard}. One
published work, now reproducible via the
\texttt{examples/attack\_defence\_survey} folder, includes a large
survey of attacks and defences against canonical datasets and models (C.
Meyers, Löfstedt, and Elmroth 2023). Another work analysed the run-time
requirements of attacks against a particular model before and after
retraining against those attacks (Charles Meyers et al. 2024)
(reproducible via \texttt{examples/retraining}). The next paper
formalised a method for estimating the time-to-failure of a given model
against a suite of attacks and introduce a metric that quantifies the
ratio of attack and training cost (Meyers et al. 2023) (reproducible via
\texttt{examples/survival\_heuristic}). Furthermore, a not yet published
work uses this time-to-failure model as a mechanism for analysing the
cost efficacy of various hardware choices in the context of adversarial
attacks (reproducible via \texttt{examples/power}) (C. Meyers et al.
2024). Another work exploits the tooling to train a custom model that is
designed to run client-side by using compression algorithms to measure
the distance between text (reproducible via
\texttt{examples/compression}).

@@

Since ML projects can exploit specialized hardware such as multi-core
processors or GPUs, and often rely on clusters of machines for
large-scale data processing, it was necessary to enable parallel and
distributed experiment execution and model optimization. By leveraging
the \texttt{hydra} configuration framework, \texttt{deckard}
automatically supports optimization libraries like \texttt{nevergrad}
(Bennet et al. 2021), \texttt{Adaptive\ Experimentation} (A. Developers
2025), and \texttt{optuna} (Akiba et al. 2019), making the software
modular and extensible. Additionally, experiments can be managed using a
variety of popular job schedulers, including \texttt{Ray} (Moritz et al.
2018), \texttt{Redis\ Queue} (Stamps 2025), and \texttt{slurm} (Yoo,
Jette, and Grondona 2003) for distributed jobs or \texttt{joblib} (Joblib
Developers 2025) for jobs on a single machine.

By using a declarative design, a given set of experiments can be
specified once and executed seamlessly across different back-ends without
modification to the underlying codebase. This makes \texttt{deckard}
both adaptable and scalable, suitable for use on personal laptops,
multi-node servers, or large-scale, high-performance clusters. When
configured appropriately, experiment batches can be parallelized,
enabling massive parameter sweeps, ensemble evaluations, or adversarial
robustness tests to be executed in parallel---reducing turnaround time
while maintaining strong guarantees on reproducibility and auditability.
The design of the presented software prioritizes clarity and
maintainability by capturing each experimental configuration as a YAML
artifact, making both successful and failed runs equally traceable and
shareable. This approach transforms experiment tracking from an
afterthought into a first-class component of the trustworthy ML
workflow.
